```{r echo=FALSE, render=FLASE}
library(tidyverse)
library(ggh4x)
library(ggbeeswarm)

source("experiment_data.r")
source("metrics.r")
```


# Any vs Except vs Max-only pregrouping settings comparisons
```{r}

load_data <- function() {
    data <- read_report("anyexcept_tud") |> 
        mutate(
            edition = factor(edition, ordered = T,
                             levels = c("CE3","CE4","CE10","CE11","CE14","CE17","CE18","CE23","CE39","CE42","CE45")
                ),
        )
}

prepare_data <- function(data, editions) {
    
    data_longer <- data |>
        
        filter(edition %in% editions) |> 
        
        pivot_longer(
            cols = starts_with("profile_"),
            names_pattern = "profile_(.+)_(.+)",
            names_to = c("grouping_class", "satisfaction"),
            values_to = "profile"
        ) |>
        
        mutate(
            across(contains(c("together", "count")), \(x) (NULL)),
            
            worst_rank = profile |> map(worst_obtained_rank) |> unlist(),
            
            grouping_class = factor(grouping_class, ordered = T,
                                    levels = c("solo", "max", "small", "submax"),
                                    labels = c("'Solo' students",
                                               "'Maximal pregroups' students",
                                               "'Small pregroups' students",
                                               "'Submax pregroups' students")
                             ),
            
            scenario = factor(scenario, ordered = T,
                              levels = c("MAX", "EXCEPT", "ANY")
                       ),
            
            satisfaction = factor(satisfaction, ordered = T,
                       levels = c("sat", "unsat"),
                       labels = c("together", "not together"))
        ) |>
        
        # CE, max/small/submax, sat/unsat
        group_by( edition ) |>
        
            # worst rank within the grouping for the relative part of 'relative'-AUPCR
            mutate( worst_rank__grp = max(worst_rank) ) |>
        
        ungroup() |>
        
        # calc relative-AUPCR (max-rank is the "worst rank within grouping")
        # if the grouping is empty (then set AUPCR to 1 - this only makes sense for unsat group)
        mutate(
            r_aupcr = map2( worst_rank__grp, profile, ~ ifelse(is.na(.y), NA, calc_aupcr(.x, 0, .y)) ) |> unlist() # 
        )
    
    return(data_longer)
} 

prepare_data_togetherness <- function(data, editions) {
    
    data_longer <- data |>
        
        filter(edition %in% editions) |>
        
        group_by(edition, mechanism, scenario) |>
            mutate(
                togetherness_max = ifelse(max_count==0, NA, max_together / max_count),
                togetherness_small = ifelse(small_count==0, NA, small_together / small_count),
                togetherness_submax = ifelse(submax_count==0, NA, submax_together / submax_count)
            ) |>
        ungroup() |>
        
        pivot_longer(
            cols = starts_with("togetherness_"),
            names_pattern = "togetherness_(.+)",
            names_to = c("grouping_class"),
            values_to = "togetherness"
        ) |>
        
        select(edition, mechanism, scenario, togetherness, grouping_class) |>
        
        mutate(
            grouping_class = factor(grouping_class, ordered = TRUE,
                                    levels = c("solo", "max", "small", "submax"),
                                    labels = c("'Solo' students",
                                               "'Maximal pregroups' students",
                                               "'Small pregroups' students",
                                               "'Submax pregroups' students")
                             ),
            
            scenario = factor(scenario, ordered = T,
                              levels = c("MAX", "EXCEPT", "ANY"),
                              labels = c("Max", "Exc", "Any")
                       ),
        )
}

prepare_data_satunsatcombined <- function(data, editions) {
    
    data_longer <- data |>
        
        filter(edition %in% editions) |> 
        
        pivot_longer(
            cols = starts_with("profile_"),
            names_pattern = "profile_(.+)_(.+)",
            names_to = c("grouping_class", "satisfaction"),
            values_to = "profile"
        ) |>
        
        mutate(
            # across(contains(c("together", "count")), \(x) (NULL)),
            
            worst_rank = profile |> map(worst_obtained_rank) |> unlist(),
            
            grouping_class = factor(grouping_class, ordered = T,
                                    levels = c("solo", "max", "small", "submax"),
                                    labels = c("'Solo' students",
                                               "'Maximal pregroups' students",
                                               "'Small pregroups' students",
                                               "'Submax pregroups' students")
                             ),
            
            scenario = factor(scenario, ordered = T,
                              levels = c("MAX", "EXCEPT", "ANY"),
                              labels = c("Max", "Exc", "Any")
                       ),
            
            satisfaction = factor(satisfaction, ordered = T,
                       levels = c("sat", "unsat"),
                       labels = c("together", "not together"))
        ) |>
        
        # CE, max/small/submax, sat/unsat
        group_by( edition ) |>
        
            # worst rank within the grouping for the relative part of 'relative'-AUPCR
            mutate( worst_rank__grp = max(worst_rank) ) |>
        
        ungroup() |>
        
        group_by( edition, grouping_class, scenario, mechanism ) |>
            mutate (profile_combined = sum_profiles(profile)) |>
        ungroup() |>
                
        # because we merged together/untogether into profile_combined we need to drop rows with the other factor from results
        filter(satisfaction == "together") |>
        
        # calc relative-AUPCR (max-rank is the "worst rank within grouping")
        # if the grouping is empty (then set AUPCR to 1 - this only makes sense for unsat group)
        mutate(
            r_aupcr = map2( worst_rank__grp, profile_combined, ~ ifelse(is.na(.y), NA, calc_aupcr(.x, 0, .y)) ) |> unlist() # 
        )
    
    return(data_longer)
} 
```

```{r fig.height = 8}
plot_raupcr_for_editions <- function(editions) {
    
    data <- load_data()

    color_scale_ce <- function() {   
        # Taken from the 12 color scheme at http://tsitsul.in/blog/coloropt/
        palette_11 <- c('#b80058', '#ebac23', '#008cf9', '#00bbad', '#006e00', '#d163e6',
                        '#b24502', '#ff9287', '#5954d6', '#00c6f8', '#878500', '#00a76c')
        mapping <- setNames(palette_11, levels(data$edition))
        ggplot2::scale_color_manual(values = mapping, limits = force)
    }
    
    if (is.null(editions)) {
        editions <- unique(data$edition)
    }

    # data_longer <- prepare_data(editions)
    data_longer <- prepare_data_satunsatcombined(data, editions)

    p_raupcr <- ggplot(data_longer,
            aes(x = interaction(scenario, mechanism),
                y = r_aupcr,
                group = interaction(scenario, mechanism)
            )
        ) +
        
        geom_boxplot(
            outlier.shape = NA, # hide outliers because all the dots are added with geom_beeswarm
            alpha = 0,
            color = "#00000020"
        ) +
        
        # geom_violin(
        #     scale = "count"
        # ) +
        
        
        geom_point(
            mapping	= aes(color = edition),
        ) +

        #rows = vars(satisfaction)
        facet_grid(cols = vars(grouping_class)) +
        
        color_scale_ce() +
        
        guides(x = guide_axis_nested(angle = 0)) + 
        
        labs(x = "Mechanism x Scenario combination",
             y = "relative-AUPCR",
             color = "Course edition"
        )
    
    # plot(p_raupcr)
    return(p_raupcr)
}
```

```{r fig.height = 5, fig.width = 10}
# Editions with all classes of pregroupings -> small, submax and max

editions_main <- c("CE4", "CE10", "CE11", "CE18")
editions_other <- c("CE23", "CE39", "CE45")

plot_main_set <- plot_raupcr_for_editions(editions_main)
plot(plot_main_set)

ggsave(plot = plot_main_set,
       filename = "plots/thesis/historical_anyexceptmax_main.png",
       device = "png", width = 300, height = 90, units = "mm", dpi = "print"
)

plot_other_set <- plot_raupcr_for_editions(editions_other)
plot(plot_other_set)

ggsave(plot = plot_other_set,
       filename = "plots/thesis/historical_anyexceptmax_other.png",
       device = "png", width = 300, height = 90, units = "mm", dpi = "print"
)


# plot_all <- plot_raupcr_for_editions(NULL)
# ggsave(plot = plot_all,
#        filename = "plots/thesis/historical_anyexceptmax_raupcr_o-groupingcombined.png",
#        device = "png", width = 300, height = 160, units = "mm", dpi = "print"
# )

```

Two questions:
    Is Any/Except a better scenario? Regardless of the mechanism (pick best)
    Is Fair better than Chia in the scenarios?l

More importantly, need to finish the viz for inclusion in the document. So:
   - rAUPCR must be tested
   - the AUPCR plots must be finalized, and:
   - the togetherness counts plotted to augment the AUPCR plots.


```{r fig.height = 4, fig.width = 10}

plot_togetherness <- function(data, editions) {
    
    ## get the data
    data_togetherness <- prepare_data_togetherness(data, editions)
    
    ## plot
    p_together <- ggplot(data_togetherness,
           aes(x = interaction(scenario, mechanism),
                y = togetherness,
                group = interaction(scenario, mechanism)
            )
        ) +
        
        ## visual aid
        geom_boxplot(
            outlier.shape = NA, # hide outliers due to adding dots via geom_jitter
            alpha = 0,
            color = "#00000020"
        ) +
        
        ## actual data points
        geom_point(
            mapping	= aes(color = edition),
        ) +
        
        facet_grid(cols = vars(grouping_class)) +
        color_scale_ce() +
        
        guides(x = "axis_nested") +
        
        labs(x = "Mechanism x Scenario combination",
             y = "Togetherness",
             color = "Course edition"
        )
}
    
p_together_main <- plot_togetherness(data, editions_main)
print(p_together)

ggsave(plot = p_together_main,
       filename = "plots/thesis/historical_anyexceptmax_together_main.png",
       device = "png", width = 300, height = 90, units = "mm", dpi = "print"
)

p_together_other <- plot_togetherness(data, editions_other)
print(p_together_other)

ggsave(plot = p_together_other,
       filename = "plots/thesis/historical_anyexceptmax_together_other.png",
       device = "png", width = 300, height = 90, units = "mm", dpi = "print"
)

```

   
   TODO: export png's and put into thesis.
   
   
## Testing the relative-AUPCR:


```{r}

scale_factors <- c(10, 3, 2.3)

# take the same profile
profile <- "3|0|2|2|0|0|1"
worst <- worst_obtained_rank(profile)

profile_parsed <- profile_histo(profile)
scaled_profile <- vector()

for (scale_factor in scale_factors) {
    
    scaled_profile <- profile_parsed * scale_factor
    scaled_profile[4] <- scaled_profile[4] + 1
    
    scaled_profile_serialized <- paste(scaled_profile, collapse  = "|")
    
    print(paste("Scale factor: ", scale_factor))
    
    print(paste("Base:     ", calc_aupcr(worst, 1, profile)))
    print(paste("Scaled   :", calc_aupcr(worst, 1, scaled_profile_serialized)))
    
    
    sum_base <- sum(profile_parsed)
    
    
    sum_scaled <- sum(scaled_profile)
    
    scaled_profile_add_norm_ser <- paste(scaled_profile / sum_scaled , collapse  = "|")
    profile_norm_ser <- paste(profile_parsed / sum_base, collapse = "|")
    
    
    print(paste("Base-norm:", calc_aupcr(worst, 1, profile_norm_ser)))
    print(paste("Scaled-nm:", calc_aupcr(worst, 1, scaled_profile_add_norm_ser)))
    
}


bbbbb <- "0.37037037037037|0|0.246913580246914|0.259259259259259|0|0|0.123456790123457"



```

Relative AUPCR results in the same value for all tested scalings of the profile. But what if the number of students change? Then tough luck?



## Project prefs generator viz for thesis

```{r fig.height=10}


prefs_data <- read_csv(file_path("\\results\\gen_prefs_export.csv"), show_col_types = FALSE) |> 
    
    mutate(
        project = paste(pref_gen_name, num_students, project, sep = "_"),
        pref_gen_name = factor(pref_gen_name, ordered = T,
                               levels = c("linear_perturbed_1", "linear_perturbed_4", "realistic", "random"),
                               labels = c("nearly identical", "mostly identical", "3-peaked", "Random")
                        ),
        num_students = factor(num_students, ordered = T,
                              levels = c(50, 200, 600),
                              labels = c("50 students", "200 students", "600 students")
                       )
    ) |>
    filter(!is.na(pref_gen_name))


p_gen_prefs <- ggplot(prefs_data) +
    geom_boxplot(aes(
        x = with(prefs_data, reorder(project, rank, median)),
        y = rank
    )) + #, color = "white", fill = "darkgreen") +
    
    facet_grid2(rows = vars(pref_gen_name), cols = vars(num_students),
               shrink = T, scales = "free", drop = T, independent = T) +
    
    # TODO: free space and constom x scale widths to improve the look of the plots
    force_panelsizes(cols = c(2, 5, 7)) +
    
    scale_x_discrete(labels = NULL, name = "Project") +
    scale_y_continuous(labels = NULL, name = "Rank")

print()

ggsave(plot = p_gen_prefs,
       filename = "plots/thesis/synth_gen_prefs.png",
       device = "png", width = 500, height = 310, units = "mm", dpi = "print"
)

```