```{r echo=FALSE, render=FLASE}
library(tidyverse)
library(ggh4x)

source("experiment_data.r")
source("metrics.r")
```


# Any vs Except vs Max-only pregrouping settings comparisons
```{r fig.height = 8}
data <- read_csv("C:\\WORK\\thesis\\aidm-optimal-groups\\reports\\anyexcept_tud.csv", show_col_types = FALSE)

data_longer <- data |>
    pivot_longer(
        cols = starts_with("profile_"),
        names_pattern = "profile_(.+)_(.+)",
        names_to = c("grouping_class", "satisfaction"),
        values_to = "profile"
    ) |>
    
    mutate(
        across(contains(c("together", "count")), \(x) (NULL)),
        
        worst_rank = profile |> map(worst_obtained_rank) |> unlist(),
        
        grouping_class = factor(grouping_class, ordered = T,
                                levels = c("solo", "max", "small", "submax"),
                                labels = c("'Solo' students",
                                           "'Maximal pregroups' students",
                                           "'Small pregroups' students",
                                           "'Submax pregroups' students")
                         ),
        
        scenario = factor(scenario, ordered = T,
                          levels = c("MAX", "EXCEPT", "ANY")
                   ),
        
        edition = factor(edition, ordered = T,
                         levels = c("CE3","CE4","CE10","CE11","CE14","CE17","CE18","CE23","CE39","CE42","CE45")
                  ),
        
        satisfaction = factor(satisfaction, ordered = T,
                   levels = c("sat", "unsat"),
                   labels = c("together", "not together"))
    ) |>
    
    # CE, max/small/submax, sat/unsat
    group_by( edition, grouping_class, satisfaction ) |>
    
        # worst rank within the grouping for the relative part of 'relative'-AUPCR
        mutate( worst_rank__grp = max(worst_rank) ) |>
    
    ungroup() |>
    
    # calc relative-AUPCR (max-rank is the "worst rank within grouping")
    # if the grouping is empty (then set AUPCR to 1 - this only makes sense for unsat group)
    mutate(
        r_aupcr = map2( worst_rank__grp, profile, ~ ifelse(is.na(.y), NA, calc_aupcr(.x, 0, .y)) ) |> unlist() # 
    )
    
# Taken from the 12 color scheme at http://tsitsul.in/blog/coloropt/
palette_11 <- c('#ebac23', '#b80058', '#008cf9', '#006e00', '#00bbad', '#d163e6', '#b24502', '#ff9287', '#5954d6', '#00c6f8', '#878500', '#00a76c')

p_raupcr <- ggplot(data_longer,
        aes(x = interaction(mechanism, scenario),
            y = r_aupcr,
            group = interaction(scenario, mechanism)
        )
    ) +
    
    geom_boxplot(
        outlier.shape = NA, # hide outliers due to adding dots via geom_jitter
        alpha = 0,
        color = "#00000020"
    ) +
    
    # geom_violin(
    #     scale = "count"
    # ) +
    geom_point(
        mapping	= aes(color = edition),
        position = position_jitterdodge(dodge.width = 0.8, jitter.width = 2)
    ) +
    
    facet_grid(rows = vars(satisfaction), cols = vars(grouping_class)) +
    scale_color_manual(values = palette_11) +
    
    guides(x = "axis_nested") +
    
    labs(x = "Mechanism x Scenario combination",
         y = "relative-AUPCR",
         color = "Course edition"
    )

plot(p_raupcr)

ggsave(plot = p_raupcr,
       filename = "plots/thesis/historical_anyexceptmax_raupcr.png",
       device = "png", width = 300, height = 160, units = "mm", dpi = "print"
)

# TODO: improve facet labels, export plot
# TODO: make plot for num together - or overlay somehow...

```

Two questions:
    Is Any/Except a better scenario? Regardless of the mechanism (pick best)
    Is Fair better than Chia in the scenarios?l

More importantly, need to finish the viz for inclusion in the document. So:
   - rAUPCR must be tested
   - the AUPCR plots must be finalized, and:
   - the togetherness counts plotted to augment the AUPCR plots.


```{r fig.height = 8}
 # togetherness

data_togetherness <- data |>
    
    pivot_longer(
        cols = ends_with("_together") | ends_with("_count"),
        names_pattern = "(.+)_(.+)",
        names_to = c("grouping_class", "status"),
        values_to = "count"
    ) |>
    
    mutate(
        status = case_when(
            status == 'together' ~ 'together',
            status == 'count' ~ 'max'
        )
    ) |> 
    
    pivot_wider(
        names_from = status, values_from = count
    ) |> 
    
    select(edition, mechanism, scenario, grouping_class, together, max) |>
    
    mutate(
        proportion_sat = ifelse(max == 0, NA, together / max),
        # Don't plot "0 unsatisfied if all are together and handle case max = 0
        proportion_unsat = ifelse(together == max, NA, 1 - (together / max) ),
        together = NULL,
        max = NULL
    ) |> 
    
    pivot_longer(
        cols = starts_with("proportion"),
        names_to = "satisfaction",
        names_prefix = "proportion_",
        values_to = "proportion"
    ) |> 
    
    mutate(
        grouping_class = factor(grouping_class, ordered = TRUE,
                                levels = c("solo", "max", "small", "submax"),
                                labels = c("'Solo' students",
                                           "'Maximal pregroups' students",
                                           "'Small pregroups' students",
                                           "'Submax pregroups' students")
                         ),
        
        scenario = factor(scenario, ordered = TRUE,
                          levels = c("MAX", "EXCEPT", "ANY")
                   ),
        
        edition = factor(edition, ordered = TRUE,
                         levels = c("CE3","CE4","CE10","CE11","CE14","CE17","CE18","CE23","CE39","CE42","CE45")
                  ),
        
        satisfaction = factor(satisfaction, ordered = T,
                              levels = c("sat", "unsat"),
                              labels = c("together", "not together")
                       )
    )

p_together <- ggplot(data_togetherness,
       aes(x = interaction(mechanism, scenario),
            y = proportion,
            group = interaction(scenario, mechanism)
        )
    ) +
    
    geom_boxplot(
        outlier.shape = NA, # hide outliers due to adding dots via geom_jitter
        alpha = 0,
        color = "#00000020"
    ) +
    
    # geom_violin(
    #     scale = "count"
    # ) +
    geom_point(
        mapping	= aes(color = edition),
        position = position_jitterdodge(dodge.width = 0.8, jitter.width = 2)
    ) +
    
    facet_grid(rows = vars(satisfaction), cols = vars(grouping_class)) +
    scale_color_manual(values = palette_11) +
    
    guides(x = "axis_nested") +
    
    labs(x = "Mechanism x Scenario combination",
         y = "Proportion students (not) together",
         color = "Course edition"
    )
    
print(p_together)
ggsave(plot = p_together,
       filename = "plots/thesis/historical_anyexceptmax_together.png",
       device = "png", width = 300, height = 160, units = "mm", dpi = "print"
)

```

   
   TODO: export png's and put into thesis.
   
   
## Testing the relative-AUPCR:


```{r}

scale_factors <- c(10, 3, 2.3)

# take the same profile
profile <- "3|0|2|2|0|0|1"
worst <- worst_obtained_rank(profile)

profile_parsed <- profile_histo(profile)
scaled_profile <- vector()

for (scale_factor in scale_factors) {
    
    scaled_profile <- profile_parsed * scale_factor
    scaled_profile[4] <- scaled_profile[4] + 1
    
    scaled_profile_serialized <- paste(scaled_profile, collapse  = "|")
    
    print(paste("Scale factor: ", scale_factor))
    
    print(paste("Base:     ", calc_aupcr(worst, 1, profile)))
    print(paste("Scaled   :", calc_aupcr(worst, 1, scaled_profile_serialized)))
    
    
    sum_base <- sum(profile_parsed)
    
    
    sum_scaled <- sum(scaled_profile)
    
    scaled_profile_add_norm_ser <- paste(scaled_profile / sum_scaled , collapse  = "|")
    profile_norm_ser <- paste(profile_parsed / sum_base, collapse = "|")
    
    
    print(paste("Base-norm:", calc_aupcr(worst, 1, profile_norm_ser)))
    print(paste("Scaled-nm:", calc_aupcr(worst, 1, scaled_profile_add_norm_ser)))
    
}


bbbbb <- "0.37037037037037|0|0.246913580246914|0.259259259259259|0|0|0.123456790123457"



```

Relative AUPCR results in the same value for all tested scalings of the profile. But what if the number of students change? Then tough luck?

